---
title: "Modeling Master Control"
author: "Josh, Joel, Corinn, Gaby"
date: "2026-02-28"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    embed-resources: true
execute:
  include: true
  eval: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: sentence
---

# Introduction
This notebook builds an initial model to help identify which MX leads are more likely to be successful. Using the cleaned dataset from the EDA phase, the goal is to better understand which lead characteristics are associated with progression to SQL, SQO, or Won, so Marketing and Sales can target higher potential leads more effectively.

### Load Data

In this section, we load the cleaned modeling-ready dataset,`leads_cleaned_final.rds` that was created during the EDA phase. This gives us a consistent starting point for modeling and allows the notebook to focus on prediction without repeating all of the earlier data cleaning steps.

```{r Load Data, include=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(rpart.plot)

leads_cleaned_final <- readRDS("data/leads_cleaned_final.rds")

#glimpse(leads_cleaned_final)
dim(leads_cleaned_final)

"solution.1" %in% names(leads_cleaned_final)
```


# Create Modeling Dataset

In this section, we prepare the final dataset that will be used for modeling. Because the business problem is focused on MX targeting, we keep only MX leads, remove fields that should not be used as predictors, and make sure the target variable is labeled clearly for interpretation.

```{r Modeling Dataset}
# Keep MX leads only and remove IDs + original outcome
mx_model_data <- leads_cleaned_final %>%
  filter(solution_rollup == "Mx") %>%
  select(-qal_id, -contact_lead_id, -next_stage_c)

# Drop leftover duplicate column if present
if ("solution.1" %in% names(mx_model_data)) {
  mx_model_data <- mx_model_data %>%
    select(-solution.1)
}

# Relabel target for readability
mx_model_data <- mx_model_data %>%
  mutate(
    next_stage_target = factor(
      next_stage_target,
      levels = c("0", "1"),
      labels = c("Not Successful", "Successful")
    )
  )

# Quick checks
dim(mx_model_data)
count(mx_model_data, next_stage_target)
"solution.1" %in% names(mx_model_data)


```

### Interpretation of Output

The modeling dataset now includes only **MX leads**, since that is the product line this project is focused on. After removing ID fields, the original outcome column, and the leftover duplicate column, the final MX modeling dataset contains 4,125 rows and 376 columns.

The target variable is also clearly imbalanced. Out of the 4,125 MX leads, 3,592 were not successful and 533 were successful, where successful means the lead progressed to **SQL, SQO, or Won**. This means only about **12.9%** of MX leads reached a successful outcome.

From a business perspective, this tells us that successful MX leads are relatively uncommon, which is consistent with what we found in the EDA. For the marketing team, this is important because it confirms that the current pool of MX leads contains many more low converting leads than high converting ones. Because of that, the goal of modeling is not just to predict the majority group, but to help identify the smaller group of leads with a higher chance of success so targeting can be improved.



# Train/Test Split

The next step is to split the MX dataset into a training set and a test set. The training set will be used to build the model, and the test set will be used later to check how well the model performs on new data. We use a 70/30 split, and we keep the success rate similar in both groups so the comparison is fair.

```{r Train and Test Split}

# Create 70/30 train/test split while keeping target balance similar
set.seed(123)

# Create split index
split_index <- createDataPartition(
  y = mx_model_data$next_stage_target,
  p = 0.70,
  list = FALSE
)

# Create training and test datasets
mx_training_data <- mx_model_data[split_index, ]
mx_test_data <- mx_model_data[-split_index, ]

# Quick checks
dim(mx_training_data)
dim(mx_test_data)

# Class balance by dataset
bind_rows(
  mx_training_data %>% mutate(dataset = "Training Set"),
  mx_test_data %>% mutate(dataset = "Test Set")
) %>%
  count(dataset, next_stage_target)
```

### Interpretation of Output

The MX dataset was successfully split into a training set with 2,889 rows and a test set with 1,236 rows, which matches the intended 70/30 split. This gives us one dataset to build the model and another separate dataset to evaluate how well the model performs on data it has not seen before.

The balance of successful and not successful leads stayed very similar in both groups. In the training set, 2,515 leads were not successful and 374 were successful, while in the test set, 1,077 leads were not successful and 159 were successful. This is important because it means both datasets reflect the same overall success pattern, giving us a fairer and more reliable way to evaluate whether the model can help identify higher-potential MX leads.

# Performance Benchmark

For this project, a useful model should perform better than random guessing and should show a meaningful ability to separate higher-potential MX leads from lower-potential ones. Because the target variable is imbalanced, ROC AUC is one of the most important benchmark metrics. As an initial guideline, a model with ROC AUC around 0.70 or higher would suggest that the model is capturing useful predictive signal.

At the same time, business usefulness also matters. Even if the model is not perfect, it can still be valuable if it helps Marketing and Sales better prioritize leads that are more likely to progress to SQL, SQO, or Won compared with the current untargeted approach

# Modeling

## Baseline Model

In this section, we build a baseline model using logistic regression. This gives us a simple starting point for predicting whether an MX lead will be successful. Starting with a baseline model is helpful because it gives us an initial benchmark that stronger models can later be compared against.

For this project, logistic regression is a good first choice because it is easier to interpret than more complex models. Since the business goal is to understand which lead characteristics, such as job title and industry, are associated with higher conversion, a baseline model helps us begin answering that question in a way that is still understandable for the business team.

```{r BaseLine Model}
# Create cross-validation folds on the training data
set.seed(123)

mx_folds <- vfold_cv(
  mx_training_data,
  v = 5,
  strata = next_stage_target
)

# Create recipe for preprocessing
mx_recipe <- recipe(next_stage_target ~ ., data = mx_training_data) %>%
  step_rm(contact_lead_title) %>%
  step_novel(all_nominal_predictors()) %>%
  step_other(all_nominal_predictors(), threshold = 0.01) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

# Define logistic regression model
mx_log_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Build workflow
mx_log_workflow <- workflow() %>%
  add_recipe(mx_recipe) %>%
  add_model(mx_log_model)

# Fit model with cross-validation
mx_log_results <- fit_resamples(
  mx_log_workflow,
  resamples = mx_folds,
  metrics = metric_set(
    yardstick::roc_auc,
    yardstick::accuracy,
    yardstick::precision,
    yardstick::recall,
    yardstick::f_meas
  )
)

# View performance metrics
collect_metrics(mx_log_results)

```


### Interpretation of Output

The baseline logistic regression model ran successfully and produced an initial set of performance metrics using 5-fold cross-validation on the training data. The model achieved an accuracy of 0.86 and a **ROC AUC of 0.75**. Because the target variable is imbalanced, ROC AUC is especially important here. A value of **0.75** suggests that the model has a reasonable ability to separate leads that are more likely to be successful from those that are less likely to be successful.

At this stage, the model is being evaluated only on the training data through cross-validation, while the test set is intentionally saved for later use with the final selected model. This helps protect the test set from influencing model development. Since the baseline ROC AUC is above the initial benchmark of **0.70**, the results suggest that the dataset contains useful predictive signal. At the same time, the warnings indicate that logistic regression may not be the most stable final model for this wide dataset, so it should be treated as a starting benchmark that later models can be compared against.

## Decision Tree 

In this section, a decision tree model is used as a second approach to predict MX lead success. Unlike logistic regression, a decision tree can capture nonlinear patterns and interactions more naturally, while also producing simpler rules that are easier to explain to a business audience.

This is especially useful for this project because the goal is not only to predict success, but also to better understand which types of leads should be prioritized. A decision tree can help translate the data into more practical decision rules around variables such as industry, enrichment quality, and title-related signals.

```{r Decision Tree}
# Create decision tree model
mx_tree_model <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()
) %>%
  set_engine("rpart") %>%
  set_mode("classification")

# Build workflow
mx_tree_workflow <- workflow() %>%
  add_recipe(mx_recipe) %>%
  add_model(mx_tree_model)

# Create tuning grid
mx_tree_grid <- grid_regular(
  cost_complexity(),
  tree_depth(),
  min_n(),
  levels = 3
)

# Tune decision tree with cross-validation
set.seed(123)

mx_tree_results <- tune_grid(
  mx_tree_workflow,
  resamples = mx_folds,
  grid = mx_tree_grid,
  metrics = metric_set(
    yardstick::roc_auc,
    yardstick::accuracy,
    yardstick::precision,
    yardstick::recall,
    yardstick::f_meas
  )
)

# Show top-performing tree models based on ROC AUC
show_best(mx_tree_results, metric = "roc_auc")

# Select best tree based on ROC AUC
best_tree <- select_best(mx_tree_results, metric = "roc_auc")
best_tree

# Finalize workflow with best parameters
final_tree_workflow <- finalize_workflow(
  mx_tree_workflow,
  best_tree
)

# Fit finalized decision tree on training data
final_tree_fit <- fit(final_tree_workflow, data = mx_training_data)
```


### Interpretation of Output
The decision tree model was tuned using cross-validation on the training data, testing different combinations of tree depth, minimum node size, and cost complexity. The best-performing tree achieved a ROC AUC of 0.82, which is higher than the baseline logistic regression ROC AUC of 0.75. This suggests that the decision tree is doing a better job of separating higher-potential MX leads from lower-potential ones.

The selected tree used a tree depth of 8 and a minimum node size of 21, which suggests that lead success depends on multiple interacting factors rather than just one or two variables alone. From a business perspective, this is encouraging because it means the model may be capturing more realistic lead patterns. It also suggests that a decision tree may be a stronger option than the baseline logistic regression for identifying which types of MX leads Marketing and Sales should prioritize.



```{r Plot Tree}
# Create a simpler tree model for interpretation
simple_tree_model <- decision_tree(
  tree_depth = 3,
  min_n = 40
) %>%
  set_engine("rpart") %>%
  set_mode("classification")

# Build workflow
simple_tree_workflow <- workflow() %>%
  add_recipe(mx_recipe) %>%
  add_model(simple_tree_model)

# Fit simpler tree
simple_tree_fit <- fit(simple_tree_workflow, data = mx_training_data)

# Extract fitted tree
simple_tree_plot_fit <- extract_fit_parsnip(simple_tree_fit)$fit

# Plot simpler tree
rpart.plot(
  simple_tree_plot_fit,
  type = 3,
  extra = 104,
  fallen.leaves = TRUE,
  roundint = FALSE
)
```

#### Interpretation of the plot

This simplified decision tree shows the first few decision rules the model is using to separate stronger MX leads from weaker ones. The first split is based on `Priority 2`, which in the data dictionary refers to a more standard-priority lead rather than a higher-intent lead such as a demo request or contact-us inquiry. In simple terms, the model is first checking whether the lead looks more routine or more engaged. That makes business sense, because lead priority can reflect how ready the prospect may be to move forward.

The next important split is `mfg_lowinfo`, which shows whether the manufacturing model information is missing or low quality. This is consistent with the EDA findings, where incomplete enrichment was associated with weaker lead quality. After that, the tree uses `solution_Manufacturing`, which is a dummy variable created from the `solution` field. This does not refer to all MX leads overall. It specifically shows whether the lead is associated with the `Manufacturing` category inside that variable. In the strongest branch of this simplified tree, leads that are **not Priority 2**, have **better manufacturing information**, and are associated with **Manufacturing** end in the most favorable success node. Overall, the plot suggests that stronger MX leads are driven by a combination of **higher engagement**, **better enrichment quality**, and **better business fit**, rather than by only one variable.

#### What the Boxes in the Tree Mean

Each box in the tree is a final prediction group, and it gives three pieces of information:

-   The **top line** shows the predicted class, such as **Successful** or **Not Successful**
-   The **middle numbers** show the estimated probabilities for each class  
    -   for example, `.34 .66` means about **34% Not Successful** and **66% Successful**
-   The **bottom percentage** shows the share of leads that fall into that box

So if a green box says **Successful / .34 .66 / 4%**, that means:

-   the model predicts that group as **Successful**
-   about **66% of leads in that group were successful**
-   and that group represents **4%** of the data in the tree

#### What We Can Infer From This Plot

This plot suggests that the model is finding some practical screening rules for MX leads. Leads that appear more engaged, have more complete manufacturing-related information, and fit the manufacturing product context seem more promising. On the other hand, leads with **Priority 2** or **low-info manufacturing fields** move early into lower-success branches, which supports the EDA finding that weaker enrichment is a negative signal.

At the same time, this plot does **not** give a clean ranking of the exact best job titles. The tree is mainly showing the strongest top-level splits, and in this simplified version, title variables do not appear in the first few branches. That does not mean titles are unimportant. It only means that, in this tree, other variables were stronger early splitters. Because of that, this plot is most helpful for understanding business rules.

## Explainable Boosting Machine (EBM)

In this section, we fit an Explainable Boosting Machine (EBM) as a third modeling approach. EBMs are a type of Generalized Additive Model with pairwise interactions (GA²M) that use cyclic gradient boosting to learn shape functions for each feature. Unlike black-box models, EBMs are fully interpretable — each feature's contribution can be visualized independently — while still achieving accuracy competitive with ensemble methods like random forests and gradient boosting.

EBMs are especially useful for this project because they can show exactly *how* each lead characteristic (industry, territory, title role, enrichment quality) relates to conversion probability, which directly supports the business goal of understanding what makes a strong MX lead.

This section uses the `ebm` R package, which wraps the Python `interpret` library via `reticulate`. The EBM section is fully self-contained — it independently loads data, joins cluster labels from the EDA phase, and creates its own train/test split using the same seed and method as the shared pipeline to ensure identical row partitions.

```{r EBM Setup, include=FALSE}
library(reticulate)
use_virtualenv("r-ebm", required = TRUE)
library(ebm)
library(patchwork)
```

```{r EBM Data Prep}
# Load cluster labels exported from EDA
cluster_labels    <- readRDS("data/cluster_labels.rds")
mx_cluster_labels <- readRDS("data/mx_cluster_labels.rds")

# Build EBM dataset independently — join cluster features, filter to Mx
ebm_data <- readRDS("data/leads_cleaned_final.rds") %>%
  left_join(cluster_labels, by = "qal_id") %>%
  left_join(mx_cluster_labels, by = "qal_id") %>%
  filter(solution_rollup == "Mx") %>%
  mutate(
    next_stage_target = factor(
      next_stage_target,
      levels = c("0", "1"),
      labels = c("Not Successful", "Successful")
    )
  ) %>%
  # Replace NAs in title_cluster with explicit level
  # (titles flagged as low-info were never clustered — EBM handles NAs
  #  internally but cannot display them in shape-function plots)
  mutate(
    title_cluster = fct_na_value_to_level(title_cluster, level = "No_Title_Info")
  ) %>%
  select(-qal_id, -contact_lead_id, -next_stage_c,
         -contact_lead_title, -solution_rollup)

# Drop leftover duplicate column if present
if ("solution.1" %in% names(ebm_data)) {
  ebm_data <- ebm_data %>% select(-solution.1)
}

# Same seed + stratified split as shared pipeline → identical row partition
set.seed(123)
ebm_split_idx <- caret::createDataPartition(
  y = ebm_data$next_stage_target, p = 0.70, list = FALSE
)
ebm_train <- ebm_data[ebm_split_idx, ]
ebm_test  <- ebm_data[-ebm_split_idx, ]

dim(ebm_train)
dim(ebm_test)
```

### Interpretation of EBM Data Prep

The EBM dataset is built independently from the shared pipeline to avoid introducing cluster-label NAs into the existing GLM and decision tree recipes. Three engineered features from the EDA clustering phase — `account_cluster`, `title_cluster`, and `mx_cluster` — are joined onto the base modeling data. For `title_cluster`, rows where the lead title was flagged as low-info (and therefore never clustered) have their NA replaced with an explicit `"No_Title_Info"` factor level so that EBM can visualize the shape function for that group.

The train/test split uses the same `set.seed(123)` and `caret::createDataPartition()` call as the shared pipeline, which guarantees identical row selection since the function depends only on the target vector and the seed. This means the EBM is evaluated on exactly the same holdout set as the GLM and decision tree, making AUC comparisons fair.

```{r EBM Fit}
# Fit Explainable Boosting Machine
mx_ebm <- ebm(next_stage_target ~ ., data = ebm_train, objective = "log_loss")
print(mx_ebm)
```

### Interpretation of EBM Fit

The `ebm()` call fits a GA²M model using cyclic gradient boosting. Each feature gets its own smooth shape function (main effect), and the model also automatically detects and fits the most important pairwise interaction terms. The `log_loss` objective is appropriate for binary classification — it optimizes the model to encourage well-calibrated predicted probabilities rather than just class labels.

The printed summary shows the number of main effects and interaction terms retained, along with the intercept. Unlike the GLM (which required dummy encoding via `step_dummy`) or the decision tree (which splits on one variable at a time), the EBM learns flexible nonlinear relationships for each feature while remaining fully interpretable.

```{r EBM Predictions}
# Generate predictions on the held-out test set
ebm_probs  <- predict(mx_ebm, newdata = ebm_test)
ebm_labels <- predict(mx_ebm, newdata = ebm_test, type = "class")

# Confusion matrix
table("Observed" = ebm_test$next_stage_target, "Predicted" = ebm_labels)
```

### Interpretation of EBM Predictions

The confusion matrix shows how the EBM classifies test-set leads compared to their actual outcomes. The key metrics to watch are:

- **True Positives (Successful predicted as Successful):** how many real successes the model catches
- **False Negatives (Successful predicted as Not Successful):** how many real successes the model misses
- **False Positives (Not Successful predicted as Successful):** leads the model incorrectly flags as promising

Because the target is imbalanced (~12.9% success rate), raw accuracy can be misleading — a model that predicts "Not Successful" for every lead would still be ~87% accurate. The calibration plot in the next section provides a better view of whether the predicted probabilities are reliable.

```{r EBM Calibration}
library(rms)
# Calibration plot: predicted probability vs observed frequency
par(las = 1)
rms::val.prob(
  ebm_probs[, 2L],
  as.integer(ebm_test$next_stage_target == "Successful"),
  cex = 0.5
)
```

### Interpretation of EBM Calibration

The calibration plot compares the model's predicted probabilities against the actual observed success rates. A well-calibrated model should have points falling close to the 45-degree diagonal line — meaning that when the model says a lead has a 20% chance of success, roughly 20% of those leads actually succeed.

Key statistics reported on the plot include the C-statistic (concordance index, equivalent to ROC AUC) and the Brier score. The C-statistic can be directly compared to the GLM's ROC AUC of 0.79 and the decision tree's ROC AUC of 0.82 to see whether the EBM provides additional predictive lift.

```{r EBM Global Explanation}
# Top 15 most important terms (main effects + interactions)
plot(mx_ebm, n_terms = 15)

# Main-effect shape function for a key feature
plot(mx_ebm, term = "acct_territory_rollup", horizontal = TRUE)

# List any detected pairwise interactions
mx_ebm$term_names_[grepl("&", mx_ebm$term_names_)]
```

### Interpretation of EBM Global Explanation

The term importance plot ranks each feature (and interaction term) by its mean absolute contribution to the predicted log-odds of success. Features at the top of the chart have the largest overall influence on whether a lead is predicted as successful or not.

The shape function plot for `acct_territory_rollup` shows exactly how each territory category shifts the prediction — for example, whether Americas-based leads get a positive boost while APAC leads get a negative one. Unlike a logistic regression coefficient (which gives a single number), the EBM shape function can show nonlinear and non-monotonic relationships.

The listed pairwise interactions reveal which feature combinations the model found to be jointly predictive beyond their individual effects. These interactions can surface insights that single-variable analyses miss — for instance, a specific industry × territory combination that converts unusually well.

```{r EBM Local Explanation}
# Local explanation for a single test-set lead
newx <- subset(ebm_test[1L, ], select = -next_stage_target)
newy <- ebm_test$next_stage_target[1L]
plot(mx_ebm, local = TRUE, X = newx, y = newy)
```

### Interpretation of EBM Local Explanation

The local explanation waterfall plot decomposes the prediction for a single lead into the contributions from each feature. Starting from the intercept (the baseline log-odds of success), each bar shows how much a specific feature pushes the prediction up (toward Successful) or down (toward Not Successful).

This is especially useful for the business team because it can answer questions like: *"Why did the model score this particular lead highly?"* or *"What would need to change about this lead for it to be predicted as successful?"* Each local explanation is fully faithful to the model — it is not an approximation — because EBMs are additive by design.

```{r EBM Save}
# Save the fitted EBM model as a Python pickle
# (EBM models use reticulate and cannot be saved with saveRDS)
reticulate::py_save_object(mx_ebm, filename = "data/mx_ebm.pkl")


colnames(predict(mx_ebm, newdata = ebm_test[1:3, ]))
```

